\LARGE \bf Mean Estimation from Adaptive One-bit Measurements

### 2017-10-09 Writeups/onebit_adaptive_Allerton_V3.tex
small changes to final manuscript

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference]{IEEEtran}      % Use this line for a4 paper
%\IEEEoverridecommandlockouts                              %

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the following constraint: the estimator can access only a single bit of each sample from this distribution. We study the squared error risk in this estimation as a function of the number of samples and one-bit measurements $n$. We consider an adaptive estimation setting where the single-bit sent at step $n$ is a function of both the new sample and the previous $n-1$ acquired bits. For this setting, we show that no estimator can attain asymptotic mean squared error smaller than $\pi/(2n)+o(n^{-1})$ times the variance. 
%
In other words, one-bit restriction increases the number of samples required for a prescribed accuracy of estimation by a factor of at least $\pi/2$ compared to the unrestricted case.
%{\color{red} This fundamental restriction on estimation in the adaptive setting implies a lower bound any distributed estimation setting from one-bit samples. }
%
In addition, we provide an explicit estimator that attains this asymptotic error, showing that, rather surprisingly, only $\pi/2$ times more samples are required in order to attain estimation performance equivalent to the unrestricted case. 
\end{abstract}

@@

Processing and estimating information from data collected at multiple physical locations is subject to communication constraints. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected through multiple sensors and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from data collected by the sensors is dictated not only by quality of observations and their numbers, but also by the available communication rate between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by communication constraints, and what are the fundamental performance limits in estimating a parameter subject to these restrictions. In this paper we answer this question in a particular simple setting: the estimation of the mean $\theta$ of a normal distribution with variance $\sigma^2$ under the constraint that only a single bit can be communicated on each sample $X_n$ from this distribution. As it turns out, the ability to share information among different samples before sending each single-bit message dramatically affects the performance in estimating $\theta$. We distinguish in particular among three different settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message which is a function of $X_1,\ldots,X_n$. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages.
 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. Since without bit constraint this risk is equals $\sigma^2/n$, we are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting, defined by the ratio between their MSE to $\sigma^2/n$ as $n$ goes to infinity when this MSE has a limit. The results in this paper focus on setting (ii), which corresponds to an estimation scenario with feedback or with multiple rounds of communication between a sensor and an estimation unit. \\% Evidently, the lower bound on the ARE we derive for setting (ii) also applies to setting (iii).

In setting (i) the estimator can evaluate the sample mean and then communicate it using $n$ bits. This strategy leads to MSE behavior of $\sigma^2/n + O(2^{-n})$. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. In this work we show that a similar result does not hold even in setting (ii): the ARE of any adaptive estimation scheme is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $1.57$ in the number of samples compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that attains this ARE. In other words, we show that the lower bound of $\pi/2$ on the asymptotic relative efficiency is tight, and that it is attained regardless of the particular realization of $\theta$ or the size of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether such efficiency is achievable (or otherwise, what is the minimal ARE) remains open.  \\

@@

We note that even through its ARE is $1$, the centralized encoding setting (i) already poses a non-trivial challenge for the design and analysis of an optimal encoding and estimation scheme. Indeed, the standard technique to encode an unknown random quantity using $n$ bits is equivalent to the design of a scalar quantizer \cite{gray1998quantization}. However, the optimal design of this quantizer depends on the distribution of its input, which is the goal of our estimation task. Therefore, a non-trivial exploration exploitation tradeoff arise in this case. In our setting, the only missing parameter is the mean, which, under setting (i), is known to the encode with uncertainty interval proportional to $\sigma/\sqrt{n}$. Therefore, while it is clear that uncertainty due to quantization decreases exponentially in the number of bits $n$, an exact expression for the MSE in this setting is not easy to derive. The situation is even more involved in the adaptive encoding of setting (ii): an encoding and estimation strategy that is globally optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$, may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $2^{n-1}$ messages, is not guaranteed to be globally optimal. Therefore, the main contribution of this paper is a lower bound on the MSE risk of any globally optimal strategy. This lower bound is obtained by showing that the Fisher information of any $n$ steps strategy is not larger than $2n/(\pi \sigma^2)$. The desired bound of $\pi \sigma^2/(2n)$ on the MSE follows from the van Trees version of the information inequality \cite{gill1995applications}. Finally, we show that an estimator that attains asymptotic MSE of $\sigma^2\pi/(2n)$ is obtained as a special case of \cite[Thm. 4]{polyak1992acceleration}. In addition to these two results, we also derive the one-step optimal strategy and demonstrates numerically that  the MSE under this strategy converge to $\pi \sigma^2/(2n)$. A proof for this converges remains elusive. \\

Our results imply that even under coarse quantization constraints, it is possible to achieve MSE in parametric estimation within only a relatively small penalty compared to the unconstrained estimator. A possible clue for this non-intuitive result is obtained from drawing the connection between our setting and the remote multiterminal source coding problem, also known as the CEO problem \cite{berger1996ceo, viswanathan1997quadratic, oohama1998rate, prabhakaran2004rate}. This connection, explained in details in Section~\ref{sec:ceo}, immediately leads to a lower bound of $4/3$ on the ARE in the distributed encoding of setting (iii). While this lower bound provides no new information compared to the lower bound of $\pi /2$ we derive later for setting (ii), it shows that the distributed nature of the problem is not a limiting factor in achieving MSE close to optimal even under one-bit quantization of each sample. \\

Setting (ii) and (iii) were considered in 
\cite{zhang2013information} under the more general assumption of $m$ machines each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ with which each machine can communicate. Specializing their result to our setting, by taking $m=n$ and $R=1$, leads to looser bounds then $\sigma^2\pi/(2n)$ for case (ii) and (iii). 
Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, although their results apply only to setting (i). The counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.
%
Estimation problem subject to the one-bit quantization constraint were also considered recently in compressed sensing \cite{boufounos20081,baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}. \\

@@

\subsubsection*{Proof of Lem.~\ref{lem:bound_intervals}}
We use induction on $n \in \mathbb N$. For the base case $n=1$ we have 
\begin{equation} \label{eq:induction_base}
\frac{  \phi^2(x)} 
{\Phi(x) \left(1 - \Phi(x) \right) },
\end{equation}
which is the weight function in the probit analysis and is known to be a strictly decreasing function of $|x|$ \cite{Samford1953}. In particular, the maximum of \eqref{eq:induction_base} is obtained at $x=0$ where it equals $2/\pi$. \par
Assume now that \eqref{eq:bound_intervals} holds for all integers up to some $n = N-1$ and consider the case $n = N$. The maximal value of \eqref{eq:bound_intervals} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 
\begin{align*}
& g(x_1,\ldots, x_N) \triangleq 2 \log \left(  \sum_{k=1}^{N} (-1)^{k+1} \phi(x_k) \right) - \\
&  \log
\left( \sum_{k=1}^N (-1)^{k+1} \Phi(x_k) \right)
-\log \left(1 -  \sum_{k=1}^N (-1)^{k+1} \Phi(x_k) \right) \\
& = 2 \log \delta_N - \log \Delta_N - \log \left(1 - \Delta_N  \right),
\end{align*}
where we denoted $\delta_N \triangleq \sum_{k=1}^{N} (-1)^{k+1} \phi(x_k)$ and $\Delta_N =  \sum_{k=1}^N (-1)^{k+1} \Phi(x_k)$. The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by
\[
\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} \phi'(x_k)}{\delta_N} -\frac{(-1)^{k+1} \phi(x_k)}{\Delta_N } + \frac{(-1)^{k+1} \phi(x_k)}{1-\Delta_N }.
\]
Using the fact that $\phi'(x) = -x \phi(x)$, we conclude that the gradient of $g$ vanishes only if
\[
x_k = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.
\]
In particular, the condition above implies $x_1 = \ldots = x_N$. If $N$ is odd then for $x_1=\ldots =x_N$ we have that the LHS of \eqref{eq:bound_intervals} equals
\[
\frac{\phi(x_1)^2}{ \Phi(x_1) (1 - \Phi(x_1))},
\]
which was shown to be not larger than $2/pi$. If $N$ is even, then for any constant $c$ the limit of the LHS of \eqref{eq:bound_intervals} as $(x_1,\ldots,x_N)\rightarrow (c,\ldots,c)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:bound_intervals} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:bound_intervals} is maximized at the borders, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches plus or minus infinity. For simplicity we only consider the cases where $x_N$ goes to minus infinity or $x_1$ goes to plus infinity (the general case where the first $m$ coordinates goes to infinity or the last $m$ to minus infinity is obtained using similar arguments). Assume first $x_N \rightarrow -\infty$. Then the LHS of \eqref{eq:bound_intervals} equals
\begin{align*}
\frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}\phi(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} \Phi(x_k) \right)\left(1- \sum_{k=1}^{N-1} (-1)^{k+1} \Phi(x_k)  \right) } ,
\end{align*}
which is smaller than $2/\pi$ by the induction hypothesis. Assume now that $x_1 \rightarrow \infty$. Then the LHS of
\eqref{eq:bound_intervals} equals
\begin{align*}
& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}\phi(x_k) \right)^2} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} \Phi(x_k) \right)\left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} \Phi(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}\phi(x'_m) \right)^2} 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} \Phi(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} \Phi(x'_{m})  \right) },
\end{align*}
where $x'_{m} = x_{m+1}$. The last expression is also smaller than $2/\pi$ by the induction hypothesis. This proves Lem.~\ref{lem:bound_intervals}. \\

@@

\section{Acknowledgments}
The work of A. Kipnis is partially supported by NSF Center for Science of Information (CSoI) under grant CCF-0939370 and NSF-BSF under grant 1609695.

>>>>

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference,9pt]{IEEEtran}      % Use this line for a4 paper
%\IEEEoverridecommandlockouts                              %

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the following constraint: the estimator can access only a single bit from each sample from this distribution. We study the squared error risk in this estimation as a function of the number of samples and one-bit measurements $n$. We consider an adaptive estimation setting where the single-bit sent at step $n$ is a function of both the new sample and the previous $n-1$ acquired bits. For this setting, we show that no estimator can attain asymptotic mean squared error smaller than $\pi/(2n)+O(n^{-2})$ times the variance. 
%
In other words, one-bit restriction increases the number of samples required for a prescribed accuracy of estimation by a factor of at least $\pi/2$ compared to the unrestricted case.
%{\color{red} This fundamental restriction on estimation in the adaptive setting implies a lower bound any distributed estimation setting from one-bit samples. }
%
In addition, we provide an explicit estimator that attains this asymptotic error, showing that, rather surprisingly, only $\pi/2$ times more samples are required in order to attain estimation performance equivalent to the unrestricted case. 
\end{abstract}

@@

Estimation performance of information from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected through multiple sensors and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number, but also by available rate of communication between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by communication constraints, and what are the fundamental performance limits in estimating a parameter subject to these restrictions. In this paper we answer this question in a particular setting: the estimation of the mean $\theta$ of a normal distribution with variance $\sigma^2$ under the constraint that only a single bit can be communicated on each sample $X_n$ from this distribution. As it turns out, the ability to share information among different samples before committing on each single-bit message dramatically affects the performance in estimating $\theta$. We therefore distinguish in among three settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message which is a function of $X_1,\ldots,X_n$. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages.
 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. We are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting relative to the MSE attained by the sample mean, which is the minimax estimator in estimating without one-bit constraint and its MSE decreases as $\sigma^2/n+O(n^{-2})$. \\

In setting (i) the estimator can evaluate the sample mean and then communicate it using $n$ bits. This strategy leads to MSE behavior of $\sigma^2/n + O(2^{-n})$. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. In this work we show that a similar result does not hold even in setting (ii): the ARE of any adaptive estimation scheme is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $1.57$ in the number of samples compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that attains this ARE. In other words, we show that the lower bound of $\pi/2$ on the ARE is tight, and that it is attained regardless of the particular realization of $\theta$ or the radius of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether this efficiency is achievable (or otherwise, what is the minimal ARE) remains open. 
\\

The lower bound of $\pi/2$ on the ARE, i.e., a lower bound of $\sigma^2\pi/(2n) +O(n^{-2})$ on the MSE, is obtained by showing that the Fisher information of any $n$ adaptive messages is not larger than $2n/(\pi \sigma^2)$. From here, the desired bound on the MSE follows from the van Trees version of the information inequality \cite{gill1995applications}. Finally, we show that an estimator that attains asymptotic MSE of $\sigma^2\pi/(2n)$ is obtained as a special case of \cite[Thm. 4]{polyak1992acceleration}. In addition to these two results, we also derive the one-step optimal strategy in which the message sent at step $i$th is designed to minimize the MSE given this message and the previous $i-1$ messages. Furthermore, we demonstrates numerically that the MSE under this strategy converge to $\pi \sigma^2/(2n)$. %A proof for this converges remains elusive. \\

\subsection*{Background and Related Works}

@@

We note that even though its ARE is $1$, the centralized encoding setting (i) already poses a non-trivial challenge for the design and analysis of an optimal encoding and estimation scheme. Indeed, the standard technique to encode an unknown random quantity using $n$ bits is equivalent to the design of a scalar quantizer \cite{gray1998quantization}. However, the optimal design of this quantizer depends on the distribution of its input, which is the goal of our estimation problem and hence its exact value is unknown. As a result, a non-trivial exploration exploitation tradeoff arises in this case. In our setting, the only missing parameter is the mean, which, under setting (i), is known to the encoder with uncertainty interval proportional to $\sigma/\sqrt{n}$. Therefore, while it is clear that uncertainty due to quantization decreases exponentially in the number of bits $n$ leading to ARE $1$, an exact expression for the MSE in this setting is not easy to derive. The situation is even more involved in the adaptive encoding of setting (ii): an encoding and estimation strategy that is optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$, may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $2^{n-1}$ messages, is not guaranteed to be globally optimal. Therefore, the main contribution of this paper is a lower bound on the MSE risk of any globally optimal strategy. \\

Our result of ARE of $\pi/2$ in the adaptive setting implies that even under coarse quantization constraints it is possible to achieve MSE in parametric estimation within only a relatively small penalty compared to the unconstrained estimator. A possible clue for this non-intuitive result is obtained from drawing the connection between our setting and the remote multiterminal source coding problem, also known as the CEO problem \cite{berger1996ceo, viswanathan1997quadratic, oohama1998rate, prabhakaran2004rate}. This connection, which is explained in details in Section~\ref{sec:ceo}, immediately leads to a lower bound of $4/3$ on the ARE in the distributed encoding of setting (iii). While this lower bound provides no new information compared to the lower bound of $\pi /2$ we derive later for setting (ii), it shows that the distributed nature of the problem is not a limiting factor in achieving MSE close to optimal even under one-bit quantization of each sample.\\

We also note that our settings (ii) and (iii) are special cases of 
\cite{zhang2013information} that consider adaptive and distributed estimation protocols for $m$ machines, each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ each machine uses for communication. The specialization of their result to our setting by, taking $m=n$ and $R=1$, leads to looser bounds then $\sigma^2\pi/(2n)$ for cases (ii) and (iii). 
Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, although their results apply only to setting (i). The counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.
%
Estimation problem subject to the one-bit quantization constraint were also considered recently in compressed sensing \cite{boufounos20081,baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}. \\

\subsection*{Paper Organization}

@@

\subsubsection*{Proof of Lem.~\ref{lem:bound_intervals}}
We use induction on $n \in \mathbb N$. For the base case $n=1$ we have 
\begin{equation} \label{eq:induction_base}
\frac{  \phi^2(x)} 
{\Phi(x) \left(1 - \Phi(x) \right) },
\end{equation}
which is the weight function in the probit analysis and is known to be a strictly decreasing function of $|x|$ \cite{Samford1953}. 
%The weight function describes the variance of the MLE in the probit model 
In particular, the maximum of \eqref{eq:induction_base} is obtained at $x=0$ where it equals $2/\pi$. \par
Assume now that \eqref{eq:bound_intervals} holds for all integers up to some $n = N-1$ and consider the case $n = N$. The maximal value of \eqref{eq:bound_intervals} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 
\begin{align*}
& g(x_1,\ldots, x_N) \triangleq 2 \log \left(  \sum_{k=1}^{N} (-1)^{k+1} \phi(x_k) \right) - \\
&  \log
\left( \sum_{k=1}^N (-1)^{k+1} \Phi(x_k) \right)
-\log \left(1 -  \sum_{k=1}^N (-1)^{k+1} \Phi(x_k) \right) \\
& = 2 \log \delta_N - \log \Delta_N - \log \left(1 - \Delta_N  \right),
\end{align*}
where we denoted $\delta_N \triangleq \sum_{k=1}^{N} (-1)^{k+1} \phi(x_k)$ and $\Delta_N =  \sum_{k=1}^N (-1)^{k+1} \Phi(x_k)$. The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by
\[
\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} \phi'(x_k)}{\delta_N} -\frac{(-1)^{k+1} \phi(x_k)}{\Delta_N } + \frac{(-1)^{k+1} \phi(x_k)}{1-\Delta_N }.
\]
Using the fact that $\phi'(x) = -x \phi(x)$, we conclude that the gradient of $g$ vanishes only if
\[
x_k = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.
\]
In particular, the condition above implies $x_1 = \ldots = x_N$. If $N$ is odd then for $x_1=\ldots =x_N$ we have that the LHS of \eqref{eq:bound_intervals} equals
\[
\frac{\phi(x_1)^2}{ \Phi(x_1) (1 - \Phi(x_1))},
\]
which was shown to be not larger than $2/pi$. If $N$ is even, then for any constant $c$ the limit of the LHS of \eqref{eq:bound_intervals} as $(x_1,\ldots,x_N)\rightarrow (c,\ldots,c)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:bound_intervals} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:bound_intervals} is maximized at the borders, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches plus or minus infinity. For simplicity we only consider the cases where $x_N$ goes to minus infinity or $x_1$ goes to plus infinity (the general case where the first $m$ coordinates goes to infinity or the last $m$ to minus infinity is obtained using similar arguments). Assume first $x_N \rightarrow -\infty$. Then the LHS of \eqref{eq:bound_intervals} equals
\begin{align*}
\frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}\phi(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} \Phi(x_k) \right)\left(1- \sum_{k=1}^{N-1} (-1)^{k+1} \Phi(x_k)  \right) } ,
\end{align*}
which is smaller than $2/\pi$ by the induction hypothesis. Assume now that $x_1 \rightarrow \infty$. Then the LHS of
\eqref{eq:bound_intervals} equals
\begin{align*}
& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}\phi(x_k) \right)^2} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} \Phi(x_k) \right)\left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} \Phi(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}\phi(x'_m) \right)^2} 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} \Phi(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} \Phi(x'_{m})  \right) },
\end{align*}
where $x'_{m} = x_{m+1}$. The last expression is also smaller than $2/\pi$ by the induction hypothesis. This proves Lem.~\ref{lem:bound_intervals}. \\

@@

%\section{Acknowledgments}
%The work of A. Kipnis is partially supported by NSF Center for Science of Information (CSoI) under grant CCF-0939370 and NSF-BSF under grant 1609695.
